---
title: "Revenue optimization"
author: "Rob Trangucci & Jonah Gabry"
date: "January 9, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rstan)
library(dplyr)
library(reshape2)
library(bayesplot)
library(ggplot2)
```

## Problem set-up

Let's say that we're a data scientist at a large media company. We are told that
we have the power to set prices for all streaming movies at whatever we want in
order to maximize revenue. How should we go about learning how to maximize
revenue for each movie? Our data might look like the number of movies sold per
day, $y_{b,t}$ at a price $p_{b,t}$.

Well, one way we could do this is to use the data we have to estimate the demand curve for the 
products. The decision problem we want to solve is the following:

\begin{align*}
\arg\max_{p \in P} \mathbb{E}_q[q(p) \, p]
\end{align*}

Because we have complete control over price, the random variable contributing to the expectation
is quantity. That's what we'll be modeling for these movies, the quantity as a function of price.

The data is in a file called \texttt{movie_data.RDS}. Let's load the data and see what the structure
is:

```{r load-data}
movie_data <- readRDS('movie_data.RDS')
head(movie_data)
```

We've got the number of units sold, the title, the price (in US dollars), the release year and the 
IMDB rating.

```{r describe-data}
N_titles <- length(unique(movie_data$title))
N_titles
```

We're working with 20 titles. We should first want to learn whether price is
related to quantity for our movies.

We'll start with a simple model and build it up slowly. 

\begin{align*}
y_{b,t} & \sim \text{Poisson}(\mu_{b,t}) \\
\mu_{b,t} & = \alpha + \beta \, p_{b,t}
\end{align*}

Let's prep the data. First, because we're concerned about comparing our
predictions for certain products, and we'd like to do the comparison in Stan, we
should convert our titles to integers.

```{r prep-data}
t <- max(table(movie_data$title))
J <- length(unique(movie_data$title))
movie_data <- movie_data %>%
  mutate(title_fac = as.factor(title),
         ttl_idx = as.integer(title_fac),
         ids = rep(1:t,J),
         wk_idx = lubridate::isoweek(date),
         mo_idx = lubridate::month(date))
```

```{r stan-data}
stan_dat_simple <- with(movie_data, list(qty_sld = units_sold,
                                         price = price,
                                         N = length(price),
                                         J = J,
                                         t = t,
                                         M = length(unique(mo_idx)),
                                         W = length(unique(wk_idx)),
                                         ttl_idx = ttl_idx,
                                         wk_idx = wk_idx,
                                         mo_idx = mo_idx))
```

```{r comppoisdgp, cache=T, results="hide", message=FALSE}
comp_dgp_simple <- stan_model('simple_poisson_regression_dgp.stan')
```

```{r runpoisdgp}
fitted_model_dgp <- sampling(comp_dgp_simple, data = stan_dat_simple, chains = 1, cores = 1, iter = 1, algorithm='Fixed_param')
samps_dgp <- rstan::extract(fitted_model_dgp)
```

```{r fake_stan_dat}
stan_dat_fake <- with(movie_data, list(qty_sld = samps_dgp$y_gen[1,],
                                             price = price,
                                             N = length(price),
                                             J = J,
                                             t = t,
                                             M = length(unique(mo_idx)),
                                             W = length(unique(wk_idx)),
                                             ttl_idx = ttl_idx,
                                             wk_idx = wk_idx,
                                            mo_idx = mo_idx))

```

```{r comppois, cache=T, results="hide", message=FALSE}
comp_model_pois <- stan_model('simple_poisson_regression.stan')
```

```{r runpoisoverfake}
fitted_model_pois <- sampling(comp_model_pois, data = stan_dat_fake, chains = 4, cores = 4)
samps_pois <- rstan::extract(fitted_model_pois)
```

```{r}
print(fitted_model_pois,pars=c('alpha','beta'))
```

```{r}
hist(samps_pois$beta)
```

```{r runpois}
fitted_model_pois <- sampling(comp_model_pois, data = stan_dat_simple, chains = 4, cores = 4)
samps_pois <- rstan::extract(fitted_model_pois)
```

Let's look at our predictions vs. the data.

```{r ppc-full}
bayesplot::ppc_dens_overlay(stan_dat_simple$qty_sld, 
                            samps_pois$pp_y[1:200,]) + xlim(c(0,50))
```

It appears that our data generating process that we've encoded in our model doesn't match the
data we've been given. Are there other statistics we can look at to determine what's missing
from the model?

We can look at the errors in the means to see whether we're at least modeling the means well
at the movie level.

```{r ppc-group_means}
bayesplot::ppc_stat_grouped(stan_dat_simple$qty_sld, 
                            samps_pois$pp_y[1:200,], group = stan_dat_simple$ttl_idx, stat = 'mean')
```

In order to capture the movie-level means, let's add some more parameters to the generative
model.

\begin{align*}
y_{b,t} & \sim \text{Poisson}(\mu_{b,t}) \\
\mu_{b,t} & = \alpha_b + \beta \, p_{b,t} \\
\alpha_b & \sim \text{Normal}(\alpha, \sigma_{\alpha})
\end{align*}

```{r comp-pois-hier, cache=T, results="hide", message=FALSE}
comp_model_pois_hier <- stan_model('hier_poisson_regression.stan')
```

```{r run-pois-hier}
fitted_model_pois_hier <- sampling(comp_model_pois_hier, data = stan_dat_simple, chains = 4, cores = 4)
samps_pois_hier <- rstan::extract(fitted_model_pois_hier)
```

```{r ppc-full-hier}
bayesplot::ppc_dens_overlay(stan_dat_simple$qty_sld, 
                            samps_pois_hier$pp_y[1:200,]) + xlim(c(0,50))
```

```{r ppc-group_means-hier}
bayesplot::ppc_stat_grouped(stan_dat_simple$qty_sld, 
                            samps_pois_hier$pp_y[1:200,], group = stan_dat_simple$ttl_idx, stat = 'mean')
```

Ok, so these plots look \emph{much} better. There are still some things that
look off about the chart though. Maybe adding a hierarchy to the price variable
will help.

\begin{align*}
y_{b,t} & \sim \text{Poisson}(\mu_{b,t}) \\
\mu_{b,t} & = \alpha_b + \beta_t \, p_{b,t} \\
\alpha_b & \sim \text{Normal}(\alpha, \sigma_{\alpha})
\beta_b & \sim \text{Normal}(\beta, \sigma_{\beta})
\end{align*}

```{r comp-pois-hier-slopes, cache=T, results="hide", message=FALSE}
comp_model_pois_hier_slopes <- stan_model('hier_slopes_poisson_regression.stan')
```

```{r run-pois-hier-slopes}
fitted_model_pois_hier_slopes <- sampling(comp_model_pois_hier_slopes, data = stan_dat_simple, chains = 4, cores = 4)
samps_pois_hier_slopes <- rstan::extract(fitted_model_pois_hier_slopes)
```

```{r ppc-full-hier-slopes}
bayesplot::ppc_dens_overlay(stan_dat_simple$qty_sld, 
                            samps_pois_hier_slopes$pp_y[1:200,]) + xlim(c(0,100))
```

```{r ppc-group_max-hier-slopes-mean-by-mo}
bayesplot::ppc_stat_grouped(stan_dat_simple$qty_sld, 
                            samps_pois_hier_slopes$pp_y[1:200,], group = stan_dat_simple$mo_idx, stat = 'max')
```

We might be missing information in month

```{r comp-pois-hier-mos, cache=T, results="hide", message=FALSE}
comp_model_pois_hier_mos <- stan_model('hier_slopes_poisson_regression_mos.stan')
```

```{r run-pois-hier-slopes-mos}
fitted_model_pois_hier_mos <- sampling(comp_model_pois_hier_mos, data = stan_dat_simple, chains = 4, cores = 4)
samps_pois_hier_mos <- rstan::extract(fitted_model_pois_hier_mos)
```

```{r ppc-full-hier-mos}
bayesplot::ppc_dens_overlay(stan_dat_simple$qty_sld, 
                            samps_pois_hier_mos$pp_y[1:200,]) + xlim(c(0,100))
```

## Revenue forecasts

Let's modify the Stan program to generate revenue forecasts at different prices in the
last week of data.

We want to predict quantity at each price, and then calculate revenue for each price.

```{r comp-rev, cache=T, results="hide", message=FALSE}
comp_rev <- stan_model('hier_slopes_poisson_regression_mos_predict.stan')
```

```{r run-pois-hier-rev}
rev_model <- sampling(comp_rev, data = stan_dat_simple, chains = 4, cores = 4)
samps_rev <- rstan::extract(rev_model)
```

```{r rev-curves}
movie_map <- unique(movie_data[,c('title','ttl_idx')])
median_rev <- apply(samps_rev$hypo_rev,c(2,3),median)
lower_rev <- apply(samps_rev$hypo_rev,c(2,3),quantile,0.05)
upper_rev <- apply(samps_rev$hypo_rev,c(2,3),quantile,0.95)
rev_df <- data.frame(rev = as.vector(median_rev), lower = as.vector(lower_rev),
                     upper = as.vector(upper_rev),
                     price = rep(1:10,20),
                     ttl_idx = as.vector(sapply(1:20,rep,10)))
rev_df <- rev_df %>% left_join(movie_map, by = 'ttl_idx')
ggplot(data = rev_df, aes(x = price, y = rev)) + geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill="grey70") + facet_wrap(~ title)
```

